---
title: "NRPS analysis"
author: "Edmond Berne"
date: "2024-08-15"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r loading_packages, include=FALSE}
# .libPaths("/home/edmond/R/x86_64-pc-linux-gnu-library/4.3") # path to the R libraries (personnal environnement)
library(phyloseq)
library(ampvis2)
library(ggplot2)
library(vegan)
library(dplyr)
library(tidyr)
library(readxl)
library(rprojroot)
library(ggpubr)
library(VennDiagram)
library(plotly)
library(ggbeeswarm)
# format the html output
library(kableExtra)
```


```{r loading_data, include=FALSE}

df_feature = read.csv("ressources/data/grouped-feature-table.tsv", sep = "\t")
df_tax = read.csv("ressources/data/amplicon_annotation.csv", sep = "\t")
df_meta = read.csv("ressources/data/Metadata.csv")


df_feature = df_feature %>%
  tibble::column_to_rownames("ASV")

df_tax = df_tax %>% 
  tibble::column_to_rownames("ASV")

df_meta = df_meta %>% 
  tibble::column_to_rownames("Sample_ID")

rownames(df_meta) = gsub("-", ".", rownames(df_meta)) # replace - by . in the sample names to be compatible with the feature table

otu_mat = as.matrix(df_feature)
tax_mat = as.matrix(df_tax)
OTU = otu_table(otu_mat, taxa_are_rows = TRUE)
TAX = tax_table(tax_mat)
samples = sample_data(df_meta)
physeq = phyloseq(OTU, TAX, samples)
saveRDS(physeq, "ressources/RDS/physeq_raw.rds")
physeq@sam_data$WEEK = ifelse(physeq@sam_data$WEEK == "2weeks", "week 2", 
                              ifelse(physeq@sam_data$WEEK == "4weeks", "week 4", "week 5"))
physeq_rhizoplane = subset_samples(physeq, Compartments == "rhizoplane") # subset the rhizoplane samples
physeq_rhizosphere = subset_samples(physeq, Compartments == "rhizosphere") # subset the rhizosphere samples
saveRDS(physeq_rhizoplane, "ressources/RDS/physeq_rhizoplane.rds")
saveRDS(physeq_rhizosphere, "ressources/RDS/physeq_rhizosphere.rds")
physeq_rhizosphere_ctrl = subset_samples(physeq_rhizosphere, Treatment == "Control") # objects to use to study the control samples only
physeq_rhizosplane_ctrl = subset_samples(physeq_rhizoplane, Treatment == "Control")
```


```{r rarefaction curves, include=TRUE}
Rarecurve = function(data, col="WEEK", facet = "Treatment", type) { 
  cat("Creating rarefaction curve for ", type, " samples\n")
  p1 = amp_rarecurve(data, color_by = col, stepsize = 100, facet_by = facet)
  p1 = p1 + theme(axis.text = element_text(size = 14), 
             axis.title = element_text(size = 20), 
             legend.text = element_text(size = 18), 
             legend.title = element_text(size = 20)) +
    theme(axis.text.x = element_text(angle = 0, hjust = 0.5, size = 14),
    strip.text = element_text(size = 20)) + # Agrandir le texte du titre de chaque facette + # Mettre les graduations à l'horizontale et les agrandir
    labs(y = "Number of ASVs")+
    geom_line(linewidth = 1.5)
  cat("Saving the plot\n")
  ggsave(paste0("output/rarefaction_curve_control_", type ,".png"), width = 10, height = 10)
  return(p1)
}
update_week = function(amp_data){
  amp_data$metadata$WEEK = ifelse(amp_data$metadata$WEEK == "2weeks", "week 2",
                                  ifelse(amp_data$metadata$WEEK == "4weeks", "week 4", "week 5"))
  return(amp_data)
}
```

```{r rarefaction curves_load, include=TRUE}
amp_rhizosphere = amp_load(physeq_rhizosphere_ctrl) # creation of the ampvis2 object
amp_rhizosphere = update_week(amp_rhizosphere)
amp_rhizoplane = amp_load(physeq_rhizosplane_ctrl)
amp_rhizoplane = update_week(amp_rhizoplane)
```

```
### Rarefaction curves
- Rhizoplane
```{r rarecurve_rhizoplane, include = TRUE}
(plot_rhizoplane = Rarecurve(amp_rhizoplane, "WEEK", "Treatment", "rhizoplane"))
```
42000 looks good for the rhizoplane samples. We will use this value for the rarefaction of the rhizoplane samples

- Rhizosphere
```{r rarecurve_rhizosphere, include = TRUE}
plot_rhizosphere = Rarecurve(amp_rhizosphere, "WEEK", "Treatment", "rhizosphere")
plot_rhizosphere
```
40000 looks good for the rhizosphere samples. We will use this value for the rarefaction of the rhizosphere samples

```{r rarefaction process, include = FALSE}

rarefaction = function(physeq_obj, type, size){
    rarefied_physeq_list = list() # a list to store the rarefied phyloseq objects
    for (i in 1:100) {
      rarefied_physeq = rarefy_even_depth(physeq_obj, sample.size = size, rngseed = i)
      rarefied_physeq_list[[i]] = rarefied_physeq
      cat("Loop number: \r", i, "out of 100")
    }
    otu_tables = lapply(rarefied_physeq_list, function(x) otu_table(x)) # regroup all the otu_tables in a list

    cat("All the rarefied phyloseq objects have been created")
    cat("Converting otu_tables to dataframes...")

    for (i in 1:length(otu_tables)) { # convert all the otu_tables to dataframes
      otu_tables[[i]] = as.data.frame(otu_tables[[i]])
      otu_tables[[i]]$ASV = rownames(otu_tables[[i]])
      cat("Loop number: \r", i, "out of 100")
    }

    cat("Dataframes created")
    cat("Concatenating dataframes...")

    df_rare = as.data.frame(otu_tables[[1]]) # create a dataframe with the first otu_table for later
    for (i in 2:length(otu_tables)) { # Concatenate all the dataframes vertically
      df_rare = rbind(df_rare, as.data.frame(otu_tables[[i]]))
      cat("Loop number: \r", i, "out of 100")
    }

    cat("Rarefied dataframe created\n")
    cat("Calculating mean and median...\n")

    # Regroup rows by NAME and calculate the mean and median
    df_rare_mean = df_rare %>% group_by(ASV) %>% summarise(across(everything(), list(mean = mean)))
    df_rare_median = df_rare %>% group_by(ASV) %>% summarise(across(everything(), list(median = median)))

    names(df_rare_mean) = sub("_mean", "", names(df_rare_mean)) # Remove mean and median to column names
    names(df_rare_median) = sub("_median", "", names(df_rare_median))

    df_rare_mean = df_rare_mean %>%
      tibble::column_to_rownames("ASV")

    df_rare_median = df_rare_median %>%
      tibble::column_to_rownames("ASV")

    df_rare_mean = otu_table(df_rare_mean, taxa_are_rows = TRUE)
    df_rare_median = otu_table(df_rare_median, taxa_are_rows = TRUE)
    Rarefied_physeq_mean = phyloseq(df_rare_mean, TAX, samples) # create another phyloseq object with rarefied data as OTU table
    Rarefied_physeq_median = phyloseq(df_rare_median, TAX, samples)

    otu_data = round(otu_table(Rarefied_physeq_mean))
    # In case where we have float number in mean (for richness), we round the number to have an integer
    Rarefied_physeq_mean_rounded = phyloseq(otu_data, TAX, samples)

    cat("Saving the object to RDS format...\n")
    saveRDS(Rarefied_physeq_mean_rounded, file = paste0("output/Rarefied_physeq_mean_rounded_", type, ".rds"))
    saveRDS(Rarefied_physeq_median, file = paste0("output/Rarefied_physeq_median_", type ,".rds"))
    saveRDS(df_rare_mean, file = paste0("output/df_rare_mean_", type ,".rds"))
}
cat("Starting the rarefaction process for the rhizoplane samples\n")
rarefaction(physeq_rhizosplane_ctrl, "rhizoplane", size = 40000)
cat("Starting the rarefaction process for the rhizosphere samples\n")
rarefaction(physeq_rhizosphere_ctrl, "rhizosphere", size = 40000)
min(colSums(physeq_rhizosplane_ctrl@otu_table))
```

```{r loading the rarefied phyloseq objects, include=FALSE}
Rarefied_physeq_mean_rounded_rhizoplane = readRDS("output/Rarefied_physeq_mean_rounded_rhizoplane.rds")
Rarefied_physeq_median_rhizoplane = readRDS("output/Rarefied_physeq_median_rhizoplane.rds")
df_rare_mean_rhizoplane = readRDS("output/df_rare_mean_rhizoplane.rds")

Rarefied_physeq_mean_rounded_rhizosphere = readRDS("output/Rarefied_physeq_mean_rounded_rhizosphere.rds")
Rarefied_physeq_median_rhizosphere = readRDS("output/Rarefied_physeq_median_rhizosphere.rds")
df_rare_mean_rhizosphere = readRDS("output/df_rare_mean_rhizosphere.rds")
```

### Descriptive statistics

```{r descriptive_statistics, include=TRUE}
# setwd("/home/edmond/Edmond_Frederik_amplicon_study_NRPS/NRPS")
reads_data = read.csv("ressources/data/track_V2.txt", sep = "\t")
number_ASV = colSums(df_feature[,-1])
reads_data$ASV = number_ASV
reads_data$Treatment = df_meta$Treatment # add metadata to the reads data
reads_data$Compartments = df_meta$Compartments
table = reads_data %>% 
    filter(Treatment == "Control") %>% # remove this line to get data for all the samples
    group_by(Compartments) %>% 
    summarise(mean_reads = mean(nonchim), median_reads = median(nonchim), reads_number = sum(nonchim),
                                min_reads = min(nonchim), max_reads = max(nonchim), Cluster_sum = sum(ASV), samples = n())
table = as.data.frame(table)

table %>% # display the table with a good looking format
  kable("html", caption = "Descriptive Statistics") %>%
  kable_styling(bootstrap_options = c("striped", "hover", "responsive"), full_width = F)

```



